Reasoning About a Machine with Local Capabilities

### 2018-05-01 toplas/paper.tex
Added malloc specification assumption to the first example lemma
Added the full correctness lemmas to the appendix

In order to formulate results about \texttt{\footnotesize{f1}} and
\texttt{\footnotesize{f2}}, we need a way to observe whether the assertion
fails. To this end, we assume they have access to a flag (an address in memory).
If the assertion fails, then the flag is set to $1$ and execution halts. The correctness lemma
for \texttt{\footnotesize{f1}} then states:
\begin{lemma}
  \label{lem:correctness-f1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_\adv & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
    c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
    c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
    \end{array}
\]
where each of the capabilities have an appropriate range of authority and
pointer\footnote{These assumptions are kept intentionally vague for brevity.
  Full statements are in the technical appendix~\citep{technical_appendix}.}.
Furthermore
  \begin{itemize}
  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{\footnotesize{f1}}
  \item $\ms_\flag(\flag) = 0$
  \item $\ms_{\var{link}}$ contains $c_\adv$ and $c_\malloc$
  \item $\hs_\adv$ contains $c_\link$ and otherwise only instructions.
  \end{itemize}
  If $(\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m')$,
  then $m'(\flag) = 0$
\end{lemma}

@@

>>>>

In order to formulate results about \texttt{\footnotesize{f1}} and
\texttt{\footnotesize{f2}}, we need a way to observe whether the assertion
fails. To this end, we assume they have access to a flag (an address in memory).
If the assertion fails, then the flag is set to $1$ and execution halts. The correctness lemma
for \texttt{\footnotesize{f1}} then states:
\begin{lemma}
  \label{lem:correctness-f1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_\adv & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
    c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
    c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
    \end{array}
\]
where each of the capabilities have an appropriate range of authority and
pointer\footnote{These assumptions are kept intentionally vague for brevity.
  Full statements are in the technical appendix~\citep{technical_appendix}.}.
Furthermore
  \begin{itemize}
  \item $c_\malloc$ satisfies the specification for malloc with
    $\iota_{\malloc,0}$
  \item $\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}$
  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{\footnotesize{f1}}
  \item $\ms_\flag(\flag) = 0$
  \item $\ms_{\var{link}}$ contains $c_\adv$ and $c_\malloc$
  \item $\hs_\adv$ contains $c_\link$ and otherwise only instructions.
  \end{itemize}
  If $(\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m')$,
  then $m'(\flag) = 0$
\end{lemma}

@@

 \subsection{Example correctness lemmas}
\begin{lemma}[Correctness lemma for \texttt{f1}] \forcenewline
  \label{lem:correctness-f1-app}
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\olf,\mathtt{1f},\mathtt{f1}) \\
    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
    m & \defeq \hs_{f1} \uplus 
        \hs_\flag \uplus                
        \ms_{\var{link}} \uplus 
        \hs_\adv \uplus 
        \ms_{\malloc} \uplus 
        \hs_{\var{frame}} 
  \end{align*}
  and
  \begin{itemize}
  \item $c_\malloc$ satisfies the specification for malloc and $\iota_{\malloc,0}$ is the region from the specification.
  \end{itemize}
  where 
  \begin{align*}
    &\dom(\hs_{f1}) = [\mathtt{f1}-\olf,\mathtt{1f}] \\
    &\dom(\hs_\flag) = [\flag,\flag] \\
    &\dom(\ms_\link) = [\link,\link+1]\\
    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
  \end{align*}
  and
  \begin{itemize}
  \item $\ms_{f1}(\mathtt{f1}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f1}(\mathtt{f1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f1}$ contains the code of $f1$.
  \item $\ms_\flag = [\flag \mapsto 0]$
  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
  \item $\hs_\adv$ contains a global read-only capability for $\hs_\link$ on its first address. The remaining cells of the memory segment only contain instructions.
  \end{itemize}
  if 
  \[
    (\reg\update{\pcreg}{c_{f1}},m) \step[n] (\halted,m'),
  \]
  then
  \[
    m'(\flag) = 0
  \]  
\end{lemma}

\begin{lemma}[Correctness lemma for \texttt{f2}]
  \label{lem:correctness-f2-app}
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-\olf,\mathtt{2f},\mathtt{f2}) \\
    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
    c_\link & \defeq ((\readonly,\glob),\link,\link+1,\link)\\
    \reg & \in \Regs \\
    m & \defeq \hs_{f2} \uplus 
        \hs_\flag \uplus                
        \ms_{\var{link}} \uplus 
        \hs_\adv \uplus 
        \ms_{\malloc} \uplus 
        \ms_{\var{stk}} \uplus
        \ms_{\var{frame}} 
  \end{align*}
  and
  \begin{itemize}
  \item $c_\malloc$ satisfies the specification for malloc and $\iota_{\malloc,0}$ is the region from the specification.
  \end{itemize}
  where 
  \begin{align*}
    &\dom(\hs_{f2}) = [\mathtt{f2}-\olf,\mathtt{2f}] \\
    &\dom(\hs_\flag) = [\flag,\flag] \\
    &\dom(\ms_\link) = [\link,\link+1]\\
    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \qquad \text{ for all $n \in \nats$}
  \end{align*}
  and
  \begin{itemize}
  \item $\ms_{f2}(\mathtt{f2}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f2}(\mathtt{f2}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f2}$ contains the code of $f2$.
  \item $\ms_\flag = [\flag \mapsto 0]$
  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
  \item $\hs_\adv(\start_\adv) = c_\link$ and $\forall \addr \in [\start_\adv+1,\addrend]\ldotp \ms_\adv(a) \in \ints$
  \end{itemize}
  if 
  \[
    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
  \]
  then
  \[
    m'(\flag) = 0
  \]  
\end{lemma}
\begin{lemma}[Correctness lemma for \texttt{f3}]
  \label{lem:correctness-f3}
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
    c_{f3} & \defeq ((\rwx,\glob),\mathtt{f3}-\olf,\mathtt{3f},\mathtt{f3}) \\
    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
    c_\link & \defeq ((\readonly,\glob),\link,\link+1,\link) \\
    \reg & \in \Regs \\
    m & \defeq \hs_{f3} \uplus 
        \hs_\flag \uplus                
        \ms_{\var{link}} \uplus 
        \hs_\adv \uplus 
        \ms_{\malloc} \uplus 
        \ms_{\var{stk}} \uplus
        \ms_{\var{frame}} 
  \end{align*}
  and
  \begin{itemize}
  \item $c_\malloc$ satisfies the specification for malloc.
  \end{itemize}
  where 
  \begin{align*}
    &\dom(\hs_{f3}) = [\mathtt{f3}-\olf,\mathtt{3f}] \\
    &\dom(\hs_\flag) = [\flag,\flag] \\
    &\dom(\ms_\link) = [\link,\link+1]\\
    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
  \end{align*}
  and
  \begin{itemize}
  \item $\ms_{f3}(\mathtt{f3}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f3}(\mathtt{f3}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f3}$ contains the code of $f3$.
  \item $\ms_\flag = [\flag \mapsto 0]$
  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
  \item $\hs_\adv(\start_\adv) = c_\link$ and all other addresses of $\ms_\adv$ contain instructions.
  \end{itemize}
  if 
  \[
    (\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
  \]
  then
  \[
    m'(\flag) = 0
  \]  
\end{lemma}

 \begin{lemma}[Correctness of $g1$]
  \label{lem:correctness-g1}
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\rwx,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
    c_{g1} & \defeq ((\entry,\glob),\mathtt{g1}-\olf,\mathtt{4f},\mathtt{g1}) \\
    c_\stk & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
    c_\link & \defeq ((\readonly,\glob),\link,\link,\link) \\
    m & \defeq \hs_{g1} \uplus 
        \ms_\flag \uplus                
        \ms_\link \uplus 
        \ms_\adv \uplus 
        \ms_\malloc \uplus 
        \ms_\stk \uplus
        \ms_{\var{frame}} 
  \end{align*}
  where 
  \begin{itemize}
  \item $c_\malloc$ satisfies the specification for malloc with $\iota_{\malloc,0}$
  \end{itemize}
  \begin{align*}
    &\dom(\hs_{g1}) = [\mathtt{g1}-\olf,\mathtt{4f}] \\
    &\dom(\hs_\flag) = [\flag,\flag] \\
    &\dom(\ms_\link) = [\link,\link]\\
    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
  \end{align*}
  and
  \begin{itemize}
  \item $\ms_{g1}(\mathtt{g1}-\olf) = ((\readonly,\glob),\link,\link,\link)$, $\ms_{g1}(\mathtt{g1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{g1}$ contains the code of $g1$ immediately followed by the code of $f4$.
  \item $\ms_\flag = [\flag \mapsto 0]$
  \item $\ms_{\var{link}} = [\link \mapsto c_\malloc]$
  \item $\hs_\adv(\start_\adv) = c_\link$ and all other addresses of $\ms_\adv$ contain instructions.
  \item $\forall a \in \dom(\ms_\stk) \ldotp \ms_\stk(a) = 0$ %This condition could be weakened to \memSat{\ms_\stk}{[\iota^\pwl(\dom(\ms_\stk))]}
  \end{itemize}
  if 
  \[
    (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step[n] (\halted,m'),
  \]
  then
  \[
    m'(\flag) = 0
  \]  
\end{lemma}
